---
title: "Distributions"
output: html_notebook
---

```{r}
library(tidyverse)
library(patchwork)
library(truncnorm)
```


# Random Numbers 

Exercises taken from here: https://evalsp21.classes.andrewheiss.com/example/random-numbers/

## Uniform Distribution

Every number is equally likely.

Functions to go with uniform distributions

```{r}
possible_answers <- c(1, 2, 3, 4, 5, 6) 
sample(possible_answers, size = 1)
sample(possible_answers, size = 3)
# Choose 10 random numbers, with replacement
sample(possible_answers, size = 10, replace = TRUE)

# Choose 8 numbers between 1 and 6, but don't replace them.
# This won't work!
sample(possible_answers, size = 8)
```


```{r}
set.seed(1234)

die <- tibble::tibble(value = sample(possible_answers,
									 size = 1000,
									 replace = TRUE))

die %>%
	count(value)


ggplot(die, aes(x = value)) + 
	geom_bar() + 
	labs(title = "1,000 rolls of a single die")

```
This looks uniform!

```{r}
set.seed(1234)

die <- tibble::tibble(value = sample(possible_answers,
									 size = 10000,
									 replace = TRUE))

die %>%
	count(value)


ggplot(die, aes(x = value)) + 
	geom_bar() + 
	labs(title = "10,000 rolls of a single die")

```
Even more so.


### runif

Random numbers between 0 -1 

```{r}
runif(5)
```
```{r}
runif(5, min = 35, max = 56)
```
```{r}
# Generate 5 people between the ages of 18 and 35
round(runif(5, min = 18, max = 35), 0)
## [1] 21 28 33 34 31

```
```{r}
set.seed(1234)
lots_of_numbers <- tibble(x = runif(5000, min = 18, max = 35))

ggplot(lots_of_numbers, aes(x = x)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 18)

```

## Normal Distributions

```{r}
set.seed(1234)

plot_data <- tibble(x = rnorm(1000, mean = 10, sd = 4))
head(plot_data)
## # A tibble: 6 x 1
##        x
##    <dbl>
## 1  5.17 
## 2 11.1  
## 3 14.3  
## 4  0.617
## 5 11.7  
## 6 12.0

ggplot(plot_data, aes(x = x)) +
  geom_histogram(binwidth = 1, boundary = 0, color = "white")
```
```{r}
set.seed(1234)

plot_data <- tibble(x = rnorm(1000, mean = 10, sd = 10))
head(plot_data)
## # A tibble: 6 x 1
##        x
##    <dbl>
## 1  -2.07
## 2  12.8 
## 3  20.8 
## 4 -13.5 
## 5  14.3 
## 6  15.1

ggplot(plot_data, aes(x = x)) +
  geom_histogram(binwidth = 1, boundary = 0, color = "white")

```

Make fake people

```{r}
set.seed(1234)

fake_people <- tibble(income = rnorm(1000, mean = 40000, sd = 15000),
                      age = rnorm(1000, mean = 25, sd = 8),
                      education = rnorm(1000, mean = 16, sd = 4))
head(fake_people)
## # A tibble: 6 x 3
##   income   age education
##    <dbl> <dbl>     <dbl>
## 1 21894. 15.4      12.1 
## 2 44161. 27.4      15.6 
## 3 56267. 12.7      15.6 
## 4  4815. 30.1      20.8 
## 5 46437. 30.6       9.38
## 6 47591.  9.75     11.8

fake_income <- ggplot(fake_people, aes(x = income)) +
  geom_histogram(binwidth = 5000, color = "white", boundary = 0) +
  labs(title = "Simulated income")

fake_age <- ggplot(fake_people, aes(x = age)) +
  geom_histogram(binwidth = 2, color = "white", boundary = 0) +
  labs(title = "Simulated age")

fake_education <- ggplot(fake_people, aes(x = education)) +
  geom_histogram(binwidth = 2, color = "white", boundary = 0) +
  labs(title = "Simulated education")

fake_income + fake_age + fake_education
```
```{r}
set.seed(1234)

plot_data <- tibble(fake_age = rnorm(1000, mean = 14, sd = 5))
head(plot_data)
## # A tibble: 6 x 1
##   fake_age
##      <dbl>
## 1     7.96
## 2    15.4 
## 3    19.4 
## 4     2.27
## 5    16.1 
## 6    16.5

ggplot(plot_data, aes(x = fake_age)) +
  geom_histogram(binwidth = 2, color = "white", boundary = 0)

```

Negative values for age does not makes sense. You can use the `rtruncnorm` function to specify a range of values. From the library `(truncnorm)`

```{r}
set.seed(1234)

plot_data <- tibble(fake_age = rtruncnorm(1000, mean = 14, sd = 5, a = 12, b = 21))
head(plot_data)
## # A tibble: 6 x 1
##   fake_age
##      <dbl>
## 1     15.4
## 2     19.4
## 3     16.1
## 4     16.5
## 5     14.3
## 6     18.8

ggplot(plot_data, aes(x = fake_age)) +
  geom_histogram(binwidth = 1, color = "white", boundary = 0)
```

### Beta Distributions

Good for left skewed distirubtions.  Good for items like percentages or proportions. 100 based exams. 

Beta distributions take two paramets

shape 1
shape 2




Beta distriubtions are good at modeling probabilities of things. shape 1 and shape 2 represent specific parts of a probability formula. 


Example Test worth 10 points. Most people get 6 points. 

6 / 6 + 4

```{r}
set.seed(1234)

plot_data <- tibble(exam_score = rbeta(1000, shape1 = 6, shape2 = 4)) %>%
  # rbeta() generates numbers between 0 and 1, so multiply everything by 10 to
  # scale up the exam scores
  mutate(exam_score = exam_score * 10)

ggplot(plot_data, aes(x = exam_score)) +
  geom_histogram(binwidth = 1, color = "white") +
  scale_x_continuous(breaks = 0:10)
```



```{r}
ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 6, shape2 = 4))

```

```{r}
ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 60, shape2 = 40))

```

```{r}

ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 9, shape2 = 1), color = "blue") +
  geom_function(fun = dbeta, args = list(shape1 = 1, shape2 = 9), color = "red")


```

Play with shapes you want to emulate. 

```{r}
ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 5, shape2 = 5), color = "blue") +
  geom_function(fun = dbeta, args = list(shape1 = 2, shape2 = 5), color = "red") +
  geom_function(fun = dbeta, args = list(shape1 = 80, shape2 = 23), color = "orange") +
  geom_function(fun = dbeta, args = list(shape1 = 13, shape2 = 17), color = "brown")

```

```{r}
set.seed(1234)

plot_data <- tibble(thing = rbeta(1000, shape1 = 2, shape2 = 5)) %>%
  mutate(thing = thing * 100)
head(plot_data)
## # A tibble: 6 x 1
##   thing
##   <dbl>
## 1 10.1 
## 2 34.5 
## 3 55.3 
## 4  2.19
## 5 38.0 
## 6 39.9

ggplot(plot_data, aes(x = thing)) +
  geom_histogram(binwidth = 2, color = "white", boundary = 0)
```


### Binomial Distributions

Outcome variables: yes/no, treated/not treated, success/fail




```{r}
set.seed(1234)

# Choose 5 random T/F values
possible_things <- c(TRUE, FALSE)
sample(possible_things, 5, replace = TRUE)
## [1] FALSE FALSE FALSE FALSE  TRUE

```

```{r}
set.seed(1234)
candidates <- c("Person 1", "Person 2")
sample(candidates, size = 1, prob = c(0.8, 0.2))
```

```{r}
set.seed(1234)
fake_elections <- tibble(winner = sample(candidates,
                                         size = 1000,
                                         prob = c(0.8, 0.2),
                                         replace = TRUE))
fake_elections %>%
  count(winner)
## # A tibble: 2 x 2
##   winner       n
##   <chr>    <int>
## 1 Person 1   792
## 2 Person 2   208

ggplot(fake_elections, aes(x = winner)) +
  geom_bar()

```

rbinom() and arguments size and prob



```{r}
set.seed(1234)

rbinom(5, size = 20, prob = 0.6)

```
```{r}
set.seed(12345)

plot_data <- tibble(thing = rbinom(2000, 1, prob = 0.6)) %>%
  # Make this a factor since it's basically a yes/no categorical variable
  mutate(thing = factor(thing))

plot_data %>%
  count(thing) %>%
  mutate(proportion = n / sum(n))
## # A tibble: 2 x 3
##   thing     n proportion
##   <fct> <int>      <dbl>
## 1 0       840      0.42 
## 2 1      1160      0.580

ggplot(plot_data, aes(x = thing)) +
  geom_bar()
```

## Poisson distribution

rpois()
lambda mysterious. 

"hat’s a Poisson process: a bunch of independent random events that combine into grouped events"


```{r}



```

